### 2017-10-12 writing/paper/sections/coincidences.tex
Some changes from Henning

Coincidences among small mesh patterns have been considered by
\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
a closure result and one worked out special case to fully classify coincidences among mesh patterns
of length 2.

Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie, \(\av{\lambda} = \av{\gamma}\). Equivalently we can say that they have the same set of
\emph{containers}, \ie, \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a classical pattern of length 3
and \(p\) is a mesh pattern of length 2 to establish sufficient conditions for two such
sets to be coincident. The classical pattern \(\pi\) will be fixed and called
the \emph{dominating pattern}. Given such a dominating pattern \(\pi\) we will write \(q_1 \coinc{\pi} q_2\) if for the mesh patterns, \(q_1\) and \(q_2\),  the sets \(\av{p,q_1}\) and \(\av{p,q_2}\) are coincident.

Our first step is to calculate whether \(\av{p,q_1}\) and \(\av{p,q_2}\) are equal up to permutations of length 11 computationally, if they are we write \(q_1 \coinc[comp]{\pi} q_2\). If \(q_1 \coinc{\pi} q_2\), it can be seen that \(q_1 \coinc[comp]{\pi} q_2\), and therefore the \(\coinc{\pi}\)-equivalence classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.

We then form three Propositions called ``Dominating Pattern Rules'', if the First Dominating
Pattern Rule shows that \(q_1 \coinc{\pi} q_2\) then we write \(q_1 \coinc[(1)]{\pi} q_2\).
If a combination of the First and Second Dominating Pattern Rules show that \(\av{p,q_1}= \av{p,q_2}\) then we write \(q_1 \coinc[(2)]{\pi} q_2\); similarly we write \(q_1 \coinc[(3)]{\pi} q_2\) if a combination of all three rules shows coincidences.

>>>>

Coincidences among small mesh patterns have been considered by
\textcite{DBLP:journals/corr/ClaessonTU14}, in which the authors use the Simultaneous Shading Lemma,
a closure result and one special case to fully classify coincidences among mesh patterns
of length 2.

Recall that two patterns \(\lambda\) and \(\gamma\) are considered \emph{coincident} if the set of permutations
that avoid \(\lambda\) is the same as the set of permutations that avoid \(\gamma\),
\ie, \(\av{\lambda} = \av{\gamma}\). This is equivalent to \(\lambad\) and \(\gamma\) being contained in
the same set of permutations, \ie, \(\cont{\lambda} = \cont{\gamma}\).

We will consider the avoidance sets \(\av{\pi,p}\) where \(\pi\) is a fixed
classical pattern of length 3 and \(p\) is a mesh pattern of length 2 that
varies. The classical pattern \(\pi\) will be called the \emph{dominating
pattern}. Given such a dominating pattern \(\pi\) we will write \(p_1
\coinc{\pi} p_2\) if for the mesh patterns, \(p_1\) and \(p_2\),  the sets
\(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are equal.

Our first step is to calculate whether \(\av{\pi,p_1}\) and \(\av{\pi,p_2}\) are
equal up to permutations of length 11 computationally, if they are we write
\(p_1 \coinc[comp]{\pi} p_2\). The equivalence relation \(\coinc{\pi}\) is a
refinement of the equivalence relation \(\coinc[comp]{\pi}\), and therefore the
\(\coinc{\pi}\)-equivalence classes form partitions of the
\(\coinc[comp]{\pi}\)-coincidence classes.

If \(p_1 \coinc{\pi} p_2\), it can be seen that
\(p_1 \coinc[comp]{\pi} p_2\), and therefore the \(\coinc{\pi}\)-equivalence
classes form partitions of the \(\coinc[comp]{\pi}\)-coincidence classes.

We then prove three propositions called ``Dominating Pattern Rules'', if the
First Dominating Pattern Rule shows that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(1)]{\pi} p_2\). If a combination of the First and Second
Dominating Pattern Rules show that \(p_1 \coinc{\pi} p_2\) then we write
\(p_1 \coinc[(2)]{\pi} p_2\); similarly we write \(p_1 \coinc[(3)]{\pi} p_2\) if
a combination of all three rules shows coincidences. These three equivalence
relations are successive refinements of \(\coinc{\pi}\). Therefore if we find
that \eg, \(\coinc[(1)]{\pi} = \coinc[comp]{\pi}\) then \(\coinc[(1)]{\pi} = \coinc{\pi}\).

### 2017-10-12 writing/paper/sections/introduction.tex
Some changes from Henning

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. This original concept---a subsequence of
symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions.
\textcite{babstein2000} considered \emph{vincular} patterns (also known as
\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
look at classes of patterns where entries can also be required to be consecutive in
value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
patterns were studied by \textcite{MR2264071} to establish necessary conditions
for a Schubert variety to be Gorenstein. These definitions are subsumed under
the definition of \emph{mesh patterns}, introduced by
\textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for
certain permutation statistics.

>>>>

The study of permutation patterns began as a result of Knuth's statements on
stack sorting in \emph{The Art of Computer Programming}~\cite[p.~243,
Ex.~5,6]{Knuth:1997:ACP:260999}. The original concept---a subsequence of
symbols having a particular relative order, now known as classical
patterns---has been expanded to a variety of definitions.
\textcite{babstein2000} considered \emph{vincular} patterns (also known as
\emph{generalised} or \emph{dashed} patterns) where two adjacent entries in the
pattern can be required to be adjacent in the permutation. \textcite{MR2652101}
look at classes of patterns where entries can also be required to be consecutive in
value, these are called \emph{bivincular} patterns. \emph{Bruhat-restricted}
patterns were studied by \textcite{MR2264071} to establish necessary conditions
for a Schubert variety to be Gorenstein. These definitions are subsumed under
the definition of \emph{mesh patterns}, introduced by
\textcite{journals/combinatorics/BrandenC11} to capture explicit expansions for
certain permutation statistics.

### 2017-10-12 writing/paper/sections/meshpatterns.tex
Final comments from Henning
Some changes from Henning

A \emph{permutation} is a bijection from the set \(\nrange{n} = \setrange{1}{n}\) to itself. The
set of all such bijections is denoted \(\mathfrak{S}_n\) and has \(n!\) elements. We can
denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotso,\pi(n)}\). The
set \(\mathfrak{S}_0\) has exactly one element, the empty permutation \(\varepsilon\).

\begin{definition}{(Order isomorphism.)}
    Two strings \(\alpha_1\alpha_2\dotsm\alpha_n\) and
    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
    if they share the same relative order, \ie, \(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

The definition of order isomorphism allows us to give the meaning of containment for classical
permutation patterns.
\begin{definition}
    A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted}
    \(\sigma \preceq \pi\)) if there is some sequence \( i_1,i_2,\dotsm{},i_k\)
    such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If \(\pi\) does not contain \(\sigma\), we say that
    \(\pi\) \emph{avoids} \(\sigma\). If this is the case the sequence
    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of \(\sigma\)
    in \(\pi\). In this context \(\sigma\) is called a (\emph{classical})
    \emph{permutation pattern}.
\end{definition}

@@

Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in\nrange{n}\}\) such that there are order-preserving
injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\) satisfying the following two conditions.

\noindent Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
  \begin{enumerate}[i.]
    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\).
  \end{enumerate}\saveenum
  Define \(R_{ij} = [\alpha(i)+1,\alpha(i+1)-1]\times[\beta(j)+1,\beta(j+1)-1]\)
   for \(i,j\in[0,k]\) where \(\alpha(0)=\beta(0)=0\) and \(\alpha(k+1)=\beta(k+1)=n+1\). Then
   the second condition is
  \begin{enumerate}[i.]\resetenum
  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\).
\end{enumerate}
We call \(R_{ij}\) the \emph{region corresponding to} \(\boks{i,j}\).

\begin{example}
    The pattern \(p=\mperm{2,1,3}{\{(0,1),(0,2),(1,0),(1,1),(2,1),(2,2)\}}=
    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }\) is contained in
    \(\pi = \perm{3,4,2,1,5}\).
\end{example}

Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
are not contained in the regions corresponding to the shaded boxes in \(p\).
\begin{equation*}
    \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                               1/0,1/1,1/2,
                               2/0,2/1,2/2,
                               3/2,3/3,3/4,
                               4/2,4/3,4/4}
\end{equation*}
The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\).

We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
contained in the mesh set of \(q\).

@@

\begin{example}
    The pattern \(p=\mperm{2,1,3}{\{(0,1),(0,2),(1,0),(2,2)\}}=
    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 2/2 }\) is contained in the pattern
    \(r = \textpattern{}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                               1/0,1/1,1/2,1/3,1/4,
                               2/1,2/2,2/3,2/4,
                               3/2,3/4,
                               4/2,4/3,4/4} \) as a subpattern.
\end{example}

The highlighted points form an occurrence of \(p\) in \(r\)
\begin{equation*}
    \pattern{1,2,5}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                               1/0,1/1,1/2,1/3,1/4,
                               2/1,2/2,2/3,2/4,
                               3/2,3/4,
                               4/2,4/3,4/4}
\end{equation*}
The permutation \(\perm{4,2,3,1,5}\) also contains \(p\) in the usual sense.

>>>>

A \emph{permutation} is a bijection from the set \(\nrange{n} = \setrange{1}{n}\) to itself. The
set of all such bijections is denoted \(\mathfrak{S}_n\) and has \(n!\) elements. We can
denote an individual permutation \(\pi\in\mathfrak{S}_n\) in \emph{one-line notation} by writing the
entries of the permutation in order, therefore \(\pi = \perm{\pi(1),\pi(2),\dotsm,\pi(n)}\). The
set \(\mathfrak{S}_0\) has exactly one element, the empty permutation \(\varepsilon\).

\begin{definition}
    Two strings of integers \(\alpha_1\alpha_2\dotsm\alpha_n\) and
    \(\beta_1\beta_2\dotsm\beta_n\) are said to be \emph{order isomorphic}
    if they share the same relative order, \ie, \(\alpha_r<\alpha_s\) if and
    only if \(\beta_r<\beta_s\).
\end{definition}

The definition of order isomorphism allows us to give the meaning of containment for classical
permutation patterns.
\begin{definition}
    A permutation \(\pi\in\mathfrak{S}_n\) \emph{contains} the permutation \(\sigma\in\mathfrak{S}_k\) (\mbox{denoted}
    \(\sigma \preceq \pi\)) if there is some sequence \( i_1,i_2,\dotsc,i_k\)
    such that \(1\le i_1<i_2<\dotsm<i_k\le n\) and the sequence \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is order isomorphic to
    \(\sigma(1)\sigma(2)\dotsm\sigma(k)\). If this is the case the sequence
    \(\pi(i_1)\pi(i_2)\dotsm\pi(i_k)\) is called an \emph{occurrence} of \(\sigma\)
    in \(\pi\). If \(\pi\) does not contain \(\sigma\), we say that
    \(\pi\) \emph{avoids} \(\sigma\). In this context \(\sigma\) is called a (\emph{classical})
    \emph{permutation pattern}.
\end{definition}

@@

Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\pi\) is a subset \(\omega\) of the plot of
\(\pi, G(\pi) = \{(i,\pi(i)\mid i\in\nrange{n}\}\) such that there are order-preserving
injections \(\alpha,\beta:\nrange{k}\mapsto\nrange{n}\) satisfying the following two conditions.
Firstly, \(\omega\) is an occurrence of \(\tau\) in the classical sense
  \begin{enumerate}[i.]
    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\tau)\}\).
  \end{enumerate}\saveenum
  Define \(R_{ij} = [\alpha(i)+1,\alpha(i+1)-1]\times[\beta(j)+1,\beta(j+1)-1]\)
   for \(i,j\in[0,k]\) where \(\alpha(0)=\beta(0)=0\) and \(\alpha(k+1)=\beta(k+1)=n+1\). Then
   the second condition is
  \begin{enumerate}[i.]\resetenum
  \item if \(\boks{i,j} \in R \text{ then } R_{ij} \cap G(\pi) = \emptyset\).
\end{enumerate}
We call \(R_{ij}\) the \emph{region corresponding to} \(\boks{i,j}\).

\begin{example}
    The pattern \(p=\mperm{2,1,3}{\{\boks{0,1},\boks{0,2},\boks{1,0},\boks{1,1},\boks{2,1},\boks{2,2}\}}=
    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }\) is contained in
    \(\pi = \perm{3,4,2,1,5}\).
    Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
    an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
    are not contained in the regions corresponding to the shaded boxes in \(p\).
    \begin{equation*}
        \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                                   1/0,1/1,1/2,
                                   2/0,2/1,2/2,
                                   3/2,3/3,3/4,
                                   4/2,4/3,4/4}
    \end{equation*}
    The subsequence \(325\) is therefore an occurrence of the pattern \(p\) in \(\pi\).
\end{example}

We define containment of a mesh pattern \(p\) in another mesh pattern \(q\) as above,
with the additional condition that if \(\boks{i,j}\in R\text{ then } R_{ij}\) is
contained in the mesh set of \(q\). More formally:

@@

\begin{example}
    The pattern \(p=\mperm{2,1,3}{\{\boks{0,1},\boks{0,2},\boks{1,0},\boks{2,2}\}}=
    \textpattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 2/2 }\) is contained in the pattern
    \(q = \textpattern{}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                               1/0,1/1,1/2,1/3,1/4,
                               2/1,2/2,2/3,2/4,
                               3/2,3/4,
                               4/2,4/3,4/4} \) as a subpattern.
    The highlighted points form an occurrence of \(p\) in \(q\)
    \begin{equation*}
        \pattern{1,2,5}{4,2,3,1,5}{0/1,0/2,0/3,0/4,
                                   1/0,1/1,1/2,1/3,1/4,
                                   2/1,2/2,2/3,2/4,
                                   3/2,3/4,
                                   4/2,4/3,4/4}
    \end{equation*}
    The permutation \(\perm{4,2,3,1,5}\) also contains \(p\) in the usual sense.
\end{example}

### 2017-10-12 writing/paper/sections/wilf.tex
Some changes from Henning

\begin{definition}{(Wilf Equivalence.)}
  Two patterns \(\pi\) and \(\sigma\) are said to be \emph{Wilf-equivalent}
  if for all \(k\ge0, \size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\). Two sets of
  permutation patterns \(R\) and \(S\) are \emph{Wilf-equivalent} if for all
  \(k\ge0, \size{\av[k]{R}} = \size{\av[k]{S}}\).
\end{definition}

>>>>

\begin{definition}
  Two patterns \(\pi\) and \(\sigma\) are said to be \emph{Wilf-equivalent}
  if for all \(k\ge0, \size{\av[k]{\pi}} = \size{\av[k]{\sigma}}\). Two sets of
  permutation patterns \(R\) and \(S\) are \emph{Wilf-equivalent} if for all
  \(k\ge0, \size{\av[k]{R}} = \size{\av[k]{S}}\).
\end{definition}
